
This thesis studies statistical applications of distances on the space of laws of random probability measures. Such objects arise naturally in Bayesian nonparametrics and in modern statistical problems where the observed units are distributions rather than finite-dimensional vectors. In these settings, distances provide a way to compare the laws governing random probability measures, thereby opening the possibility of extending classical statistical questions to this more general framework. 

The thesis develops two applications of this distance-based perspective. First, it studies a hypothesis testing problem in which a distance-based approach is used to compare laws of random probability measures. Second, it investigates merging of opinions in Bayesian nonparametric models.

The rest of the introduction is organized as follows. We begin with a historical overview of distances on spaces of probability measures, emphasizing the mathematical, information-theoretic, and statistical motivations that led to their development. We then discuss the motivation for considering distances on laws of random probability measures and review related work in this direction. Finally, we describe the organization of the thesis and specific contributions to the two problems considered: a distance-based approach to hypothesis testing for laws of random probability measures, and the analysis of merging under different prior specifications in Bayesian nonparametric models.

\section{Metrics on Probability Measures: From Historical Origins to Statistical Applications}\label{section:general_intro_metrics_on_probabilities}

Historically, several distinct motivations led to the development of distances between probability measures. The first motivation is purely mathematical, with formal foundations laid following the introduction of metric spaces by Felix Hausdorff and Maurice Fr\'echet at the beginning of the 20th century. This was followed by work on the space of signed measures that led to the development of the total variation distance. In parallel, Paul L\'evy was working on defining distances between random variables and their laws. He introduced a metric for distribution functions on the real line in \cite{levy1925calcul}. In addition, in a note written for the book \cite{frechet1950recherches}, L\'evy introduced several notions of distance between random variables and between their induced laws, one of which corresponds to what is now recognized as the Wasserstein distance.

Another motivation for studying distances between probability measures arose in information theory. In \cite{kullback1951information}, the notion of information for discrimination between two distributions was introduced. The motivation was to quantify how much information observations provide for distinguishing between two hypotheses about the underlying distribution. The information-theoretic point of view was followed by a broader development of divergence measures, such as the chi-square divergence, \(f\)-divergences, and the Hellinger distance.

One of the most widely used distances on the space of probability measures is the Wasserstein distance, although it was not originally introduced as a metric on measures. Its origins can be traced back to the development of optimal transport theory. In the late 18th century, Gaspard Monge posed the question of how to move a pile of soil to fill another hole while minimizing the total transportation cost \cite{monge1781memoire}. The problem was formulated in terms of transport maps, which specify where each infinitesimal unit of mass is sent. Around 160 years later, Leonid Kantorovich introduced a relaxation of this formulation in \cite{kantorovich1942translocation}, allowing transport plans rather than deterministic maps. This relaxation led to what is now known as the Monge--Kantorovich problem:
\begin{equation}\label{equation:kantorovich_formulation}
\min_{\gamma \in \Gamma(\mu, \nu)}
\int_{\X\times \Y} c(x,y)\,\mathrm{d}\gamma(x,y),
\end{equation}
where \(\Gamma(\mu,\nu)\) denotes the set of all couplings of \(\mu\) and \(\nu\), that is, probability measures on \(\X\times\Y\) with marginals \(\mu\) and \(\nu\). This formulation was later complemented by the duality results of \cite{kantorovich1958space}.

The Monge--Kantorovich problem includes the Wasserstein distance as a special case. Indeed, when \(\X=\Y\) is equipped with a metric \(d\), and the cost is chosen as \(c(x,y)=d(x,y)^p\) for \(p\geq 1\), the corresponding optimal value defines the \(p\)-Wasserstein distance,
\[
W_p(\mu,\nu)
=
\left(
\inf_{\gamma\in\Gamma(\mu,\nu)}
\int_{\X\times\Y} d(x,y)^p\,\mathrm d\gamma(x,y)
\right)^{1/p}.
\]
However, the terminology itself has a separate historical origin. The name ``Vasershtein distance'' first appeared in \cite{dobrushin1970prescribing}, in the context of studying the existence and uniqueness of random fields, as a reference to earlier work by Leonid Vaserstein. In \cite{vaserstein1969markov}, Markov processes on denumerable products of spaces were studied, motivated by the analysis of large systems of automata, and that distance was used to study the convergence of Markov processes.

Wasserstein distances have become important tools in statistics and machine learning because they provide a geometrically meaningful discrepancy between probability measures. In particular, they are used in the theory of point estimation, where parameters are estimated by minimizing the Wasserstein distance between the empirical distribution and a model distribution \cite{basseti2006kantorovich,bernton2019parameter}. The convergence of empirical estimators based on the Wasserstein distance is studied in \cite{bobkov2019one,panaretos2019statistical}. In hypothesis testing, Wasserstein distances are used to measure departures from a null hypothesis \cite{ramdas2017wasserstein,wang2021two,hu2025two}. They are also used in goodness-of-fit testing by comparing the empirical distribution to model-based estimators \cite{hallin2021multivariate}. In computational statistics, Wasserstein distances are used to study convergence rates of MCMC algorithms \cite{qin2022wasserstein,lavenant2024convergence} and to design likelihood-free inference methods, such as Wasserstein approximate Bayesian computation, by comparing observed and simulated empirical distributions directly \cite{bernton2019abc}. In machine learning, they play a central role in generative modeling, especially in Wasserstein generative adversarial networks \cite{arjovsky2017wasserstein}, and have also been used in domain adaptation \cite{courty2016optimal}. In addition, one can use the Wasserstein distance to define the distance function between labels in classification, which leads to a distance at the level of datasets \cite{alvarezmelis2020geometric}.

Lastly, the Wasserstein distance plays an important role in quantifying the dependence between random probability measures in Bayesian nonparametric models. In many Bayesian nonparametric models, groups of observations may arise from different sources whose underlying distributions are statistically dependent. This dependence can be modelled at the level of a vector of random probability measures, where each such measure is associated with each group. For this type of problem, \cite{catalano2024wasserstein} develops a Wasserstein index of dependence for completely random measures to quantify the dependence between random probability measures.






\section{Motivation and Applications of Distances Between Laws of Random Probability Measures}
\label{section:motivation_applications_distances_on_rpms}

There are several motivations for studying distances between laws of random probability measures. In this section, we emphasize three of them: the purely mathematical analysis perspective, the Bayesian statistical perspective, and the complex data structure perspective, where probability measures themselves are treated as data objects.

From the purely mathematical analysis perspective, optimal transport provides one important source of examples. The optimal transport problem for laws on spaces of probability measures was studied in \cite{pinzi2026optimal}, where characterization results in terms of Kantorovich potentials are obtained in analogy with classical optimal transport theory, and a connection with the strict Monge problem is established. In addition, the work \cite{pinzi2025nested} studies the geometry of probability laws on Wasserstein spaces through a nested superposition principle, giving tools to analyze curves and tangent structures in this setting. Another line of work concerns establishing a Brenier theorem for laws on iterated Wasserstein spaces \cite{beiglbock2025brenier,pinzi2025totally}, and the study of gradient-flow dynamics \cite{bonet2025flowing}.

The second motivation comes from the Bayesian statistical perspective. In Bayesian statistics, priors are probability measures over possible data-generating laws. Therefore, as the framework includes probabilities over probabilities, it is natural to work with distances between such objects.

One line of work studies the merging of opinions. Here, the question is whether posterior distributions arising from different prior distributions become close as more observations are collected. Distances between laws of random probability measures provide a natural way to measure this form of posterior closeness. The paper \cite{catalano2026merging} studies this problem in the context of completely random measures. Another important example is the use of the Wasserstein over Wasserstein distance to study posterior concentration and borrowing of information in hierarchical Bayesian nonparametric models \cite{nguyen2016borrowing}. In that paper, the goal is to estimate the base probability measure in a Dirichlet process prior from the observations generated through the mixture model arising from latent random probability measures. Such a base measure is endowed with another Dirichlet prior and posterior concentration rates are provided using the Wasserstein over Wasserstein distance.

The third motivation is the complex data structure perspective. In many modern statistical problems, the observations are not scalar or vector-valued points, but probability measures. This occurs in several fields. For example, in macroeconomics, one may observe yearly income distributions or survey distributions of inflation expectations. In demographic studies, one may compare mortality distributions across countries. In biology, single-cell RNA sequencing data can be represented through distributions of cells in a state space. In machine learning, documents, or datasets may also be represented as probability measures. The use of distances on laws of random probability measures is common in machine learning. For example, in generative modelling, images can be represented as probability measures. A generative model then produces random probability measures, and the aim is to choose a model whose generated distributional objects are close to the observed ones. Wasserstein over Wasserstein can be used to define a loss function for this purpose (see \cite{dukler2019wasserstein}). Another example is the work of \cite{piening2026generalized}, where flow matching is extended to probability measures over probability measures through a Wasserstein-on-Wasserstein


\section{Organization and Contributions of the Thesis}\label{section:organization_and_contributions}

The thesis begins with a chapter that collects the background material in Bayesian nonparametrics needed for the subsequent chapters. It starts with the framework for random probability measures and exchangeability, together with de Finetti's representation theorem. It then covers completely random measures, including the role of L\'evy intensities and Laplace functionals, identifiability issues under normalization, posterior distributions, and examples. Finally, it reviews the distances on laws of random probability measures, including the Wasserstein over Wasserstein distance and the Hierarchical Integral Probability Metric, as well as the distances on laws of random measures specialized for completely random measures.

The second chapter studies a two-sample testing problem in which the observed units are probability measures. The null hypothesis is that two samples of probability measures are drawn from the same underlying law, against the alternative that the laws differ. Our contribution is to consider this problem when the samples are probability measures themselves and to propose a distance-based approach together with its theoretical properties. Specifically, we derive a threshold based on Rademacher complexity that controls the type-I error, and then introduce a permutation-based threshold that is both practical and theoretically justified, with guarantees on both the type-I error and the power of the resulting test. We note that the theoretical guarantees are established at the level of the raw observations used to estimate the underlying probability measures, which we call hierarchical samples. The method is benchmarked against existing approaches in simulations and applied to a mortality dataset.

The third chapter studies the merging of opinions in Bayesian nonparametric models: whether posterior distributions induced by two different prior distributions become close as the number of observations grows, independently of the data-generating mechanism. The framework for studying this problem is provided by \cite{catalano2026merging}, which introduces distances between laws of completely random measures and establishes identifiability results that make the comparison of posteriors meaningful. Within this framework, the known results concern merging between two Dirichlet process priors, which occurs at rate $1/n$, and merging between a generalized Gamma CRM and a Gamma CRM, where the rate is $\max\{n^{-1/(1+\sigma)}, k_n/n\}$ and merging may fail when $k_n \sim n$, where $k_n$ denotes the number of unique values in $n$ observations. The contribution of this thesis is to establish the merging rate between the Pitman--Yor process and the Dirichlet process, showing that it equals $k_n/n$ for any sequence of observations.
