\section{Threshold via a Permutation Approach}\label{section:permutation_approach}


In this section, we introduce a permutation-based framework for two-sample testing that yields improved performance in practice. As discussed earlier, the threshold derived from Rademacher complexity is overly conservative, resulting in low power for finite sample sizes. The permutation approach alleviates this issue by providing a more accurate approximation of the quantile function of the test statistic under the null hypothesis $H_0,$ while typically yielding a much smaller threshold under the alternative.

Recall that, given $Q^1,Q^2\in \probprobx$, we observe hierarchical samples $\mathbf{X}_{n,m} \sim Q^1$ and $\mathbf{Y}_{n,m} \sim Q^2$ as in Definition~\ref{def:hier_sample} and we construct $\widehat{Q}^1_{n,m}$ and $\widehat{Q}^{2}_{n,m}$, the hierarchical empirical estimators, as in Definition~\ref{definition:hier_emp_estimator}. Equivalently, given the notation $\T_m$ introduced in Definition~\ref{definition:tau_m} and the equivalent formulation~\eqref{eq:equivalent_sampling_Tm}, we observe
\[
\widehat{P}^1_{1,m}, \dots, \widehat{P}^1_{n,m} \overset{\iid}{\sim} \T_m[Q^{1}], \qquad \widehat{P}^2_{1,m}, \dots, \widehat{P}^2_{n,m} \overset{\iid}{\sim} \T_m[Q^{2}].
\]
Using these samples, we compute the distance $d(\widehat{Q}^1_{n,m}, \widehat{Q}^2_{n,m})$, where $d$ can be either WoW or HIPM. Our goal is to estimate the quantile function of the random variable $d(\widehat{Q}^1_{n,m}, \widehat{Q}^2_{n,m})$ assuming that $Q^1 = Q^2,$ and use it as a testing threshold. Ideally, such a quantile could be estimated empirically from multiple independent realizations of this random variable. In practice, however, only a single realization is available.

We circumvent this limitation by noting that under the null hypothesis $Q^1 = Q^2$, all random probability measures $\widehat{P}^k_{i,m}$ share the same distribution. Consequently, if we randomly shuffle these objects and recompute the distance, the resulting value constitutes another realization from the distribution of $d(\widehat{Q}^1_{n,m}, \widehat{Q}^2_{n,m})$. By repeating this process, we can obtain several samples of the distance between hierarchical empirical estimators. Although these samples are not independent, this dependence is typically negligible in practice when $n$ is sufficiently large.

To implement the permutation test, we first form the pooled collection of random probability measures
\[
\left\{ \widehat{P}_{1,m},\cdots, \widehat{P}_{2n,m}  \right\} :=\left\{ \widehat{P}^1_{1,m}, \dots, \widehat{P}^1_{n,m}, \widehat{P}^2_{1,m}, \dots, \widehat{P}^2_{n,m} \right\} .
\]
A random permutation of this collection is determined by a random vector $R = (R_1, \dots, R_{2n})$ drawn uniformly from all permutations of $\{1, \dots, 2n\}$. For notational brevity, we omit the explicit dependence of $R$ on $n$.

We then define the permuted hierarchical empirical estimators as
\begin{equation}\label{definition:permuted_hierarchical_estimators}
\widetilde{Q}^1_{n,m} = \frac{1}{n} \sum_{i = 1}^n \delta_{\widehat{P}_{R_i,m}} \quad \text{and} \quad \widetilde{Q}^2_{n,m} = \frac{1}{n} \sum_{i = n+1}^{2n} \delta_{\widehat{P}_{R_i,m}}.
\end{equation}
Finally, we compute the distance $d(\widetilde{Q}^1_{n,m}, \widetilde{Q}^2_{n,m}).$

Repeating this process yields approximate samples from the null distribution of $d(\widehat{Q}^1_{n,m}, \widehat{Q}^2_{n,m})$. Then, for a given significance level $\theta \in (0,1),$ we set the testing threshold equal to the empirical $(1-\theta)$-quantile of these values.

In practice, since we work with the hierarchical samples $\mathbf{X}_{n,m}$ and $\mathbf{Y}_{n,m}$, permutation is implemented by shuffling rows across the two matrices. The resulting procedure is summarized in Algorithm~\ref{algorithm::permutation_approach}.

\begin{center}
\begin{minipage}{0.85\textwidth} % Adjust 0.85 to make it narrower or wider
\begin{algorithm}[H]
\caption{Permutation test with $d \in \{\dlip,\WW\}$}
\label{algorithm::permutation_approach}
\begin{algorithmic}[1]
\Require Hierarchical samples $\mathbf{X}_{n,m},\mathbf{Y}_{n,m}$; significance level $\theta \in (0,1)$; number of permutations $n_\text{perm}$
\Ensure Decision: $1$ = reject $H_0$, $0$ = do not reject $H_0$

\State $d_{\mathrm{obs}} \gets d(\mathbf{X}_{n,m},\mathbf{Y}_{n,m})$ \Comment{Observed test statistic}
\State $\mathcal{Z} \gets \text{vcat}(\mathbf{X}_{n,m},\mathbf{Y}_{n,m})$ \Comment{Pool the $2n$ rows}
\For{$i=1,\ldots,n_\text{perm}$}
    \State $R \gets \text{permute}(\{1,\ldots,2n\})$
    \State $d_i \gets d(\mathcal{Z}_{R_{1:n}}, \mathcal{Z}_{R_{n+1:2n}})$
\EndFor

\State $r_\theta \gets \widehat{q}_{1-\theta}(d_1,\ldots,d_{n_\text{perm}})$ \Comment{Empirical $(1-\theta)$-quantile}
\State \Return $\mathbf{1}\{d_{\mathrm{obs}} > r_\theta\}$
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

We study the theoretical properties of the proposed procedure for the distance $\dlip$ by analyzing the asymptotic distributions of the observed test statistic and the corresponding threshold under the null and alternative hypotheses. To this end, we appropriately rescale the observed $\dlip$ distance and define the test statistic
\[
T_{n,m} = \sqrt{\frac{n}{2}}\dlip\left( \widehat{Q}^1_{n,m}, \widehat{Q}^2_{n,m} \right).
\]
The corresponding threshold is defined as
\begin{equation}\label{definition:permutation_threshold}
r(n,m,\theta) = \inf \left\{ t \in \R : \mathbb{P}\left(\widetilde{T}_{n,m} > t \mid \mathbf{X}_{n,m},\mathbf{Y}_{n,m} \right) \leq \theta \right\},
\end{equation}
where 
\[
\widetilde{T}_{n,m} = \sqrt{\frac{n}{2}} \, \dlip(\widetilde{Q}^1_{n,m}, \widetilde{Q}^2_{n,m}).
\]
Note that this scaling is introduced solely for theoretical analysis and is not required for the practical implementation of the algorithm.

For the distance $\dlip$ in dimension $1$, we are able to establish theoretical guarantees for the permutation test described above. The reason is that the class of measurable functions $\dstar$ used to express the distance as an IPM in this case is a Donsker class. This implies that the asymptotic distributions of both the test statistic and the permutation threshold can be expressed in terms of norms of Brownian bridges.
The permutation test can be implemented for both metrics, but the asymptotic guarantees are only available for the HIPM metric in dimension $1$. The reason is that the function class generating WoW as an IPM, namely $\mathrm{Lip}_1(\probx, \R, \W)$, is not a Donsker class in general, so the CLT-based argument used for $\dlip$ does not extend to that setting. We emphasize, however, that this does not imply that the consistency result fails for WoW; it only means that the current proof technique does not apply.
\begin{theorem}[Consistency of permutation approach]
\label{theorem:perm_consistency_dlip}
Given $\X = [a,b]$ and $Q^1, Q^2 \in \probprobx$. Let $\theta$ denote the significance level and $r(n,m, \theta)$ be the threshold from the permutation approach. Then
\begin{itemize}
    \item if $Q^1 = Q^2,$ then $$ \lim_{n,m\to \infty} \mathbb{P}\left(T_{n,m}>r(n,m, \theta)\right) = \theta.$$
    \item if $Q^1 \neq Q^2,$ then $$ \lim_{n,m\to \infty} \mathbb{P}\left(T_{n,m} > r(n,m,\theta) \right) = 1.$$
\end{itemize}
\end{theorem}
The main strategy of the proof of Theorem~\ref{theorem:perm_consistency_dlip} is to show that the test statistic and the permutation threshold converge in distribution to the same random variable under the null hypothesis; under the alternative, the test statistic diverges to infinity while the permutation threshold converges to an almost surely bounded random variable, so that the test statistic eventually exceeds the threshold. Since the proof relies on several auxiliary results, we defer it to Appendix~\ref{section:proofs_for_permutation_section}.


























